Scattering from compact objects: \\ Regge poles and the Complex Angular Momentum method

### 2019-12-11 rainbowCAM.tex
Some typos are coorected, the captions of the plots (5-6-7) are modified, the location of plots are rearranged and my email sheffield restored (i find it class)

\documentclass[aps,prd,longbibliography,reprint,twocolumn,amsmath,amssymb,amsfonts,showpacs,superscriptaddress]{revtex4-1}%reprint

@@

\renewcommand{\arraystretch}{1.3} % espace entre les lignes de la table
%\setlength{\tabcolsep}{.3cm} % espace entre les colonnes de la table

@@

\author{Mohamed \surname{Ould~El~Hadj}}
\email{med.ouldelhadj@gmail.com}

@@

Change !

@@

Leins, Nollert and Soffel \cite{Leins:1993zz} have developed a method for calculating the polar QNM frequencies of spherically symmetric spacetimes. Benhar, Berti and Ferarri (BBF) have extended this method to the axial sector. Here we shall present a generalisation the BBF method for axial \emph{and} scalar QNM frequencies ($s=2$ and $s=0$ respectively). The method is equally valid, \textit{mutatis mutandis}, for finding Regge poles $\{\lambda_n(\omega)= \ell_n +1/2\}$ of $ S_{\ell}(\omega)$.  The solution of the Regge-Wheeler equation is written in a power-series form as follows:
%
\begin{eqnarray}\label{Power_series}
% \nonumber % Remove numbering (before each equation)
 \phi_{\omega_\ell}(r)&=&\left(\frac{r}{2M}-1\right)^{i 2M\omega} \,e^{i\omega r} \sum_{n,=0}^{+\infty}a_n\left(1-\frac{b}{r}\right)^n \nonumber \\
                      &=& e^{i\omega r_*(r)}\sum_{n,=0}^{+\infty}a_n\left(1-\frac{b}{r}\right)^n
\end{eqnarray}
where $r = b$ is some point outside the star. By substituting \eqref{Power_series} in the Regge-Wheeler equation, it can be shown that the coefficients $a_n$ satisfy a four-term recurrence relation of the form:
\begin{equation}\label{Recurrence_4_terms}
\alpha_n a_{n+1} + \beta_n a_{n} +\gamma_{n} a_{n-1} +\delta_{n} a_{n-2}  = 0, \quad \forall n\geq 2
\end{equation}
where
\begin{subequations}
\begin{eqnarray}\label{Coeffs_3_termes}
 && \alpha_n = n (n+1)\left(1-\frac{2M}{b}\right)   \\
 && \beta_n  = 2i\omega b n + 3\left(\frac{2M}{b}\right)n^2-2n^2  \\
 && \gamma_n = \left(1-\frac{6M}{b}\right)n(n-1)-\frac{2M\beta_s}{b}-\ell(\ell+1)  \\
 && \delta_n = \left(\frac{2M}{b}\right)\left(n(n-2)+\beta_s\right)
\end{eqnarray}
\end{subequations}
To determine the initial conditions $a_0$ and $a_1$, we impose the continuity of $\phi_{\omega_\ell}(r)$ and its derivative in $r = b$, according to Eq~\eqref{Power_series}, we obtain:
\begin{eqnarray}\label{Initial_Conds}
 && a_0 = e^{i\omega r_*(b)\phi_{\omega_\ell}(b)}  \\
 && a_1 = b e^{-i\omega r_*(b)}\left(\frac{-i\omega b}{b-2M}\phi_{\omega_\ell}(b)+ \frac{d}{dr}\phi_{\omega_\ell}(r)\Big{|}_{r=b}\right)
\end{eqnarray}
Here, to obtain $\phi_{\omega_\ell}(b)$ and its derivative in $r=b$, we need to solve numerically the Eq.~\eqref{H_Radial_equation} inside the star and to continue the solution outside, up to $r=b$, by solving the Regge-wheeler equation.

@@

\begin{figure}[htb]
\centering
 \includegraphics[scale=0.50]{RP_R_6_2Mw_3_16_s_0_s_2}
\caption{\label{RP_R6_approx_2Mw_3_6_s_1} The Regge poles $\lambda_n(\omega)$ for the scalar field (empty markers) and for the axial gravitational waves (filled markers). %We assume $2M =1$.
}
\label{fig:rp1}
\end{figure}

@@

 \begin{figure}[htb]
\centering
 \includegraphics[scale=0.50]{RP_R_2_dot_26_2Mw_3_6_s_0_s_2}
\caption{\label{RP_R226_approx_2Mw_3_6_s_1}  The Regge poles $\lambda_n(\omega)$ for the scalar field (empty markers) and for the axial gravitational waves (filled markers). %We assume $2M =1$.
}
\label{fig:rp2}
\end{figure}

@@

\begingroup
\squeezetable
\begin{table}[htp]
\begin{threeparttable}[htp]
%\captionsetup{font=small}
\caption{\label{tab:table4} The lowest Regge poles $\lambda_{n}(\omega)$ for the axial gravitational waves. The radius of the compact bodies is $R = 2.26M$.}
\smallskip
\centering
\begin{ruledtabular}
\begin{tabular}{ccccc}
 $n$ & $2M\omega$  & $\lambda^{\text{(S-W)\tnote{1}}}_n(\omega)$  & $\lambda^{\text{(B-R)\tnote{2}}}_n(\omega)$ & $\lambda^{\text{(N-R)\tnote{3}}}_n(\omega)$
 \\ \hline
$1$  & $3$  & $5.884755+2.047850 i $  & $0.840822+1.728755 i   $  & $8.0740924+0.022432 i   $   \\
     & $6$  & $12.796673+2.200987 i  $  & $0.959779+2.193488 i   $  & $14.247709+0.699340 i  $    \\

@@

\section{Scattering and CAM theory}
In this section we calculate the scattering cross section $d\sigma / d\Omega$ by means of the Complex Angular Momentum (CAM) method, and we compare with results obtained in the standard way from a partial-wave series \cite{Dolan:2017rtj}.

@@

@@

\begin{figure*}%[h!]
 \includegraphics[scale=0.50]{Scattering_Cross_Section_R_6_2Mw_3}
\caption{\label{S_0_R_6_2Mw_3_Exact_vs_CAM} The scalar cross section of a compact bodies for $2M\omega=3$ and $R=6M$, its Regge pole approximation and the background integral contribution. The plots show the effect of including successively more Regge poles (plots 1--3). In the final plot, the background integral is added, giving a cross section which agrees well with the (regularized) partial-wave sum.}
\end{figure*}

@@

\begin{figure*}%[h!]
 \includegraphics[scale=0.50]{Scattering_Cross_Section_R_6_2Mw_16}
\caption{\label{S_0_R_6_2Mw_16_Exact_vs_CAM} The scalar cross section of a compact bodies for $2M\omega=16$ and $R=6M$, its Regge pole approximation and the background integral contribution.}
\end{figure*}

@@

\begin{figure*}%[h!]
\centering
 \includegraphics[scale=0.50]{Scattering_Cross_Section_R_2-dot-26_2Mw_3}
\caption{\label{S_0_R_2-dot-26_2Mw_3_Exact_vs_CAM} The scalar cross section of a very compact bodies for $2M\omega=3$ and $R=2.26M$ and its Regge pole approximation.}
\end{figure*}

@@

\begin{figure*}%[h!]
\centering
 \includegraphics[scale=0.50]{Scattering_Cross_Section_R_2-dot-26_2Mw_6}
\caption{\label{S_0_R_2-dot-26_2Mw_6_Exact_vs_CAM} The scalar cross section of a very compact bodies for $2M\omega=6$ and $R=2.26M$ and its Regge pole approximation.}
\end{figure*}

@@

\begin{figure}%[h!]
\centering
 \includegraphics[scale=0.50]{Rainbow_Cross_Section_R_6_2Mw_16}
\caption{\label{Rainbow_Cross_Section_R_6_2Mw_16} Rainbow scattering for compact bodies for $2M\omega=16$ and $R=6M$, its Regge pole approximation and different contributions of the sum over Regge poles.}
\end{figure}

@@

\begin{figure}%[h!]
\centering
 \includegraphics[scale=0.50]{Diff_Contribution_Cross_Section_R_2-dot-26_2Mw_6}
\caption{\label{Diff_Contribution_Cross_Section_R_2-dot-26_2Mw_6} Rainbow scattering for compact bodies for $2M\omega=16$ and $R=2.26M$, its Regge pole approximation and different contributions of the sum over Regge poles.}
\end{figure}

@@

\acknowledgments
M. O. E. H. wish to thank Antoine Folacci for various discussions concerning this work. T.S.~acknowledges financial support from EPSRC. S.R.D.~acknowledges financial support from the European Union's Horizon 2020 research and innovation programme under the H2020-MSCA-RISE-2017 Grant No.~FunFiCO-777740, and from the Science and Technology Facilities Council (STFC) under Grant No.~ST/P000800/1.

\bibliography{rainbowCAM}
\bibliographystyle{apsrev4-1}

>>>>

%\documentclass[aps,prd,longbibliography,reprint,twocolumn,amsmath,amssymb,amsfonts,showpacs,superscriptaddress]{revtex4-1}%reprint
\documentclass[aps,prd,longbibliography,reprint,twocolumn,amsmath,amssymb,amsfonts,showpacs,footnote,superscriptaddress]{revtex4-1}%reprint

@@

\renewcommand{\arraystretch}{1.3} % espace entre les lignes de la table
%\setlength{\tabcolsep}{0.3pt} % espace entre les colonnes de la table

@@

\author{Mohamed \surname{Ould~El~Hadj}}\email{m.ouldelhadj@sheffield.ac.uk}

@@

@@

Leins, Nollert and Soffel \cite{Leins:1993zz} have developed a method for calculating the polar QNM frequencies of spherically symmetric spacetimes. Benhar, Berti and Ferarri (BBF) have extended this method to the axial sector. Here we shall present a generalisation the BBF method for axial \emph{and} scalar QNM frequencies ($s=2$ and $s=0$ respectively). The method is equally valid, \textit{mutatis mutandis}, for finding Regge poles $\{\lambda_n(\omega)= \ell_n +1/2\}$ of $ S_{\ell}(\omega)$.  The solution of the Regge-Wheeler equation is written in a power-series form as follows:
%
\begin{eqnarray}\label{Power_series}
% \nonumber % Remove numbering (before each equation)
 \phi_{\omega_\ell}(r)&=&\left(\frac{r}{2M}-1\right)^{i 2M\omega} \,e^{i\omega r} \sum_{n=0}^{+\infty}a_n\left(1-\frac{b}{r}\right)^n \nonumber \\
                      &=& e^{i\omega r_*(r)}\sum_{n=0}^{+\infty}a_n\left(1-\frac{b}{r}\right)^n
\end{eqnarray}
where $r = b$ is some point outside the star. By substituting \eqref{Power_series} in the Regge-Wheeler equation, it can be shown that the coefficients $a_n$ satisfy a four-term recurrence relation of the form:
\begin{equation}\label{Recurrence_4_terms}
\alpha_n a_{n+1} + \beta_n a_{n} +\gamma_{n} a_{n-1} +\delta_{n} a_{n-2}  = 0, \quad \forall n\geq 2
\end{equation}
where
\begin{subequations}
\begin{eqnarray}\label{Coeffs_3_termes}
 && \alpha_n = n (n+1)\left(1-\frac{2M}{b}\right)   \\
 && \beta_n  = 2i\omega b n + 3\left(\frac{2M}{b}\right)n^2-2n^2  \\
 && \gamma_n = \left(1-\frac{6M}{b}\right)n(n-1)-\frac{2M\beta_s}{b}-\ell(\ell+1)  \\
 && \delta_n = \left(\frac{2M}{b}\right)\left(n(n-2)+\beta_s\right)
\end{eqnarray}
\end{subequations}
To determine the initial conditions $a_0$ and $a_1$, we impose the continuity of $\phi_{\omega_\ell}(r)$ and its derivative in $r = b$, according to Eq~\eqref{Power_series}, we obtain:
\begin{eqnarray}\label{Initial_Conds}
 && a_0 = e^{i\omega r_*(b)\phi_{\omega_\ell}(b)}  \\
 && a_1 = b e^{-i\omega r_*(b)}\left(\frac{-i\omega b}{b-2M}\phi_{\omega_\ell}(b)+ \frac{d}{dr}\phi_{\omega_\ell}(r)\Big{|}_{r=b}\right)
\end{eqnarray}
Here, to obtain $\phi_{\omega_\ell}(b)$ and its derivative in $r=b$, we need to solve numerically the Eq.~\eqref{H_Radial_equation} inside the star and to continue the solution outside, up to $r=b$, by solving the Regge-wheeler equation.

@@

\begin{figure}[htb]
\centering
 \includegraphics[scale=0.50]{RP_R_6_2Mw_3_16_s_0_s_2}
\caption{\label{RP_R6_approx_2Mw_3_6_s_1} The Regge poles $\lambda_n(\omega)$ for the scalar field (unfilled markers) and for the axial gravitational waves (filled markers). %We assume $2M =1$.
}
\label{fig:rp1}
\end{figure}

@@

 \begin{figure}[htb]
\centering
 \includegraphics[scale=0.50]{RP_R_2_dot_26_2Mw_3_6_s_0_s_2}
\caption{\label{RP_R226_approx_2Mw_3_6_s_1}  The Regge poles $\lambda_n(\omega)$ for the scalar field (unfilled markers) and for the axial gravitational waves (filled markers). %We assume $2M =1$.
}
\label{fig:rp2}
\end{figure}

@@

\begingroup
\squeezetable
\begin{table}[htp]
\fontsize{5.}{10}
\begin{threeparttable}[htp]
\captionsetup{font=small}
\caption{\label{tab:table4} The lowest Regge poles $\lambda_{n}(\omega)$ for the axial gravitational waves. The radius of the compact bodies is $R = 2.26M$.}
\smallskip
\centering
\begin{ruledtabular}
\begin{tabular}{ccccc}
 $n$ & $2M\omega$  & $\lambda^{\text{(S-W)\tnote{1}}}_n(\omega)$  & $\lambda^{\text{(B-R)\tnote{2}}}_n(\omega)$ & $\lambda^{\text{(N-R)\tnote{3}}}_n(\omega)$
 \\ \hline
$1$  & $3$  & $5.884755+2.047850 i $  & $0.840822+1.728755 i   $  & $8.0740924+0.022432 i   $   \\
     & $6$  & $12.796673+2.200987 i  $  & $0.959779+2.193488 i   $  & $14.247709+0.699340 i  $    \\

@@

\section{Scattering and CAM theory}
In this section we calculate the scattering cross section $d\sigma / d\Omega$ by means of the CAM method, and we compare with results obtained in the standard way from a partial-wave series \cite{Dolan:2017rtj}.

@@

\begin{figure*}[htp!]
 \includegraphics[scale=0.50]{Scattering_Cross_Section_R_6_2Mw_3}
  \vspace*{-0.3cm}
\caption{\label{S_0_R_6_2Mw_3_Exact_vs_CAM} The scalar cross section of a compact bodies for $2M\omega=3$ and $R=6M$, its Regge pole approximation and the background integral contribution. The plots show the effect of including successively more Regge poles (plots 1--3). In the final plot, the background integral is added, giving a cross section which agrees well with the (regularized) partial-wave sum.}
\end{figure*}

\begin{figure*}[htp!]
 \includegraphics[scale=0.50]{Scattering_Cross_Section_R_6_2Mw_16}
 \vspace*{-0.3cm}
\caption{\label{S_0_R_6_2Mw_16_Exact_vs_CAM} The scalar cross section of a compact bodies for $2M\omega=16$ and $R=6M$, its Regge pole approximation and the background integral contribution. The plots show the effect of including successively more Regge poles (plots 1--3). In the final plot, the background integral is added, giving a cross section which agrees well with the (regularized) partial-wave sum.}
\end{figure*}

\begin{figure*}[htp]
\centering
 \includegraphics[scale=0.50]{Scattering_Cross_Section_R_2-dot-26_2Mw_3}
  \vspace*{-0.3cm}
\caption{\label{S_0_R_2-dot-26_2Mw_3_Exact_vs_CAM} The scalar cross section of a very compact bodies for $2M\omega=3$ and $R=2.26M$ and its Regge pole approximation.  The plots show the effect of including successively more Regge poles (plots 1--3). In the final plot, The sum over Regge poles gives a cross section which agrees well with the (regularized) partial-wave sum for intermediate and large values of the scattering
angle.}
\end{figure*}

\begin{figure*}[htp]
\centering
 \includegraphics[scale=0.50]{Scattering_Cross_Section_R_2-dot-26_2Mw_6}
  \vspace*{-0.3cm}
\caption{\label{S_0_R_2-dot-26_2Mw_6_Exact_vs_CAM} The scalar cross section of a very compact bodies for $2M\omega=6$ and $R=2.26M$ and its Regge pole approximation. The plots show the effect of including successively more Regge poles (plots 1--3). In the final plot, The sum over Regge poles gives a cross section which agrees well with the (regularized) partial-wave sum for intermediate and large values of the scattering
angle.}
\end{figure*}

\begin{figure}[htp]
\centering
 \includegraphics[scale=0.50]{Rainbow_Cross_Section_R_6_2Mw_16}
  \vspace*{-0.3cm}
\caption{\label{Rainbow_Cross_Section_R_6_2Mw_16} Rainbow scattering for compact bodies for $2M\omega=16$ and $R=6M$, its Regge pole approximation and different contributions of the sum over Regge poles.}
\end{figure}

\begin{figure}[htp]
\centering
 \includegraphics[scale=0.50]{Diff_Contribution_Cross_Section_R_2-dot-26_2Mw_6}
  \vspace*{-0.3cm}
\caption{\label{Diff_Contribution_Cross_Section_R_2-dot-26_2Mw_6} Rainbow scattering for compact bodies for $2M\omega=16$ and $R=2.26M$, its Regge pole approximation and different contributions of the sum over Regge poles.}
\end{figure}

@@

@@

@@

@@

@@

@@

@@

\acknowledgments
M. O. E. H. wishes to thank Antoine Folacci for various discussions concerning this work and thanks Sam R. Dolan for his kind
invitation to the University of Sheffield. T.S.~acknowledges financial support from EPSRC. S.R.D.~acknowledges financial support from the European Union's Horizon 2020 research and innovation programme under the H2020-MSCA-RISE-2017 Grant No.~FunFiCO-777740, and from the Science and Technology Facilities Council (STFC) under Grant No.~ST/P000800/1.

\bibliography{rainbowCAM}
%\bibliographystyle{apsrev4-1}
